---
title: "cma-week3"
author: "Martine Besse"
date: "2023-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readr")
library("dplyr")
library("ggplot2")
```

## Preparation: import data

```{r}
wildschwein <- read_csv("data/wildschwein_BE_2056.csv")
wildschwein
```

Extract Sabi:

! specific timezone otherwise, just text we might have difference because of it, and issue with summer and winter time.

```{r}
#number of animals:
wildschwein$TierName %>% unique()

#extract sabi:
sabi <- wildschwein %>% filter(TierName == "Sabi") %>%
  filter(DatetimeUTC >= as.POSIXct("2015-07-01", tz = "UTC"), DatetimeUTC < as.POSIXct("2015-07-03", tz = "UTC"))
```

```{r}
#plot sabi trajectories
ggplot(sabi, aes(E,N, color = DatetimeUTC)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Step a) specify a temporal window

```{r}
#visualise:
sabi %>% head(50) %>% 
  ggplot(aes(DatetimeUTC, 1)) +
  geom_point()
```

Temporal window of 60 minutes: 4 fixes\
pos[n-2] to pos[n]\
pos[n-1] to pos[n]\
pos[n] to pos[n+1]\
pos[n] to pos[n+2]

```{r}
sabi <- sabi %>%
  mutate(
    n_plus1 = sqrt((lead(E)-E)^2 + (lead(N)-N)^2), #from n to n+1
    n_plus2 = sqrt((lead(E, 2)-E)^2 + (lead(N, 2)-N)^2), #from n to n+2
    n_minus1 = sqrt((lag(E, 1)-E)^2 + (lag(N, 1)-N)^2), #from n to n-1
    n_minus2 = sqrt((lag(E, 2)-E)^2 + (lag(N, 2)-N)^2) #from n to n-2
  )
#reminder:
#lead() = next value of E
#lag() = opposite of lead()
#numbers <- 1:10
#lead(numbers, n = 2) #discard first 2
#lag(numbers, n = 5) #discard last 5

sabi
```

## Step b) Measure the distance from every point to every otehr point within this temporal window
Important to specify na.rm in the mean function,  otherwise na values interfer.  

```{r}
sabi <- sabi |> 
  rowwise() |> #group by row with rowwise(), otherwise the mutates give one result for everything
  mutate(
    stepMean = mean(c(n_minus1, n_minus2, n_plus1, n_plus2), na.rm = TRUE) 
  ) |> 
  ungroup() #to ungroup rowwise()
```

## Step c) Remove “static points"
Determines a threshold where we consider our animal not to be moving.  

Check if we could use the mean as the threshold:  
```{r}
ggplot(sabi, aes(stepMean)) +
  geom_histogram(binwidth = 10) +
  geom_vline(xintercept = mean(sabi$stepMean, na.rm = TRUE))
```

Use the mean as a threshold:  
```{r}
sabi |> 
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE)) |> 
  ggplot(aes(E, N)) +
  geom_path() +
  geom_point(aes(color = static)) + # color of points show if the points are statics or not
  coord_fixed()
```

```{r}
# filter sabi data
sabi <- sabi |> 
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE)) 

sabi_filter <- sabi |>
    filter(!static)

#plot
sabi_filter |>
    ggplot(aes(E, N)) +
    geom_path() +
    geom_point() +
    coord_fixed() +
    theme(legend.position = "bottom")
```




https://bookdown.org/yihui/rmarkdown-cookbook/rstudio-shortcuts.html
ctrl shift m -> insert pipe
ctrl alt i -> insert chunk
ctrl shift + -> zoom
alt tiret -> <- 
ctrl alt c -> run current chunk
