---
title: "cma-week3"
author: "Martine Besse"
date: "2023-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readr")
library("dplyr")
library("ggplot2")
```

# Wildboar data

## Preparation: import data

```{r}
wildschwein <- read_csv("data/wildschwein_BE_2056.csv")
wildschwein
```

Extract Sabi:

! specific timezone otherwise, just text we might have difference because of it, and issue with summer and winter time.

```{r}
#number of animals:
wildschwein$TierName %>% unique()

#extract sabi:
sabi <- wildschwein %>% filter(TierName == "Sabi") %>%
  filter(DatetimeUTC >= as.POSIXct("2015-07-01", tz = "UTC"), DatetimeUTC < as.POSIXct("2015-07-03", tz = "UTC"))
```

```{r}
#plot sabi trajectories
ggplot(sabi, aes(E,N, color = DatetimeUTC)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```

## Step a) specify a temporal window

```{r}
#visualise:
sabi %>% head(50) %>% 
  ggplot(aes(DatetimeUTC, 1)) +
  geom_point()
```

Temporal window of 60 minutes: 4 fixes\
pos[n-2] to pos[n]\
pos[n-1] to pos[n]\
pos[n] to pos[n+1]\
pos[n] to pos[n+2]

```{r}
sabi <- sabi %>%
  mutate(
    n_plus1 = sqrt((lead(E)-E)^2 + (lead(N)-N)^2), #from n to n+1
    n_plus2 = sqrt((lead(E, 2)-E)^2 + (lead(N, 2)-N)^2), #from n to n+2
    n_minus1 = sqrt((lag(E, 1)-E)^2 + (lag(N, 1)-N)^2), #from n to n-1
    n_minus2 = sqrt((lag(E, 2)-E)^2 + (lag(N, 2)-N)^2) #from n to n-2
  )
#reminder:
#lead() = next value of E
#lag() = opposite of lead()
#numbers <- 1:10
#lead(numbers, n = 2) #discard first 2
#lag(numbers, n = 5) #discard last 5

sabi
```

## Step b) Measure the distance from every point to every otehr point within this temporal window
Important to specify na.rm in the mean function,  otherwise na values interfer.  

```{r}
sabi <- sabi |> 
  rowwise() |> #group by row with rowwise(), otherwise the mutates give one result for everything
  mutate(
    stepMean = mean(c(n_minus1, n_minus2, n_plus1, n_plus2), na.rm = TRUE) 
  ) |> 
  ungroup() #to ungroup rowwise()
```

## Step c) Remove “static points"
Determines a threshold where we consider our animal not to be moving.  

Check if we could use the mean as the threshold:  
```{r}
ggplot(sabi, aes(stepMean)) +
  geom_histogram(binwidth = 10) +
  geom_vline(xintercept = mean(sabi$stepMean, na.rm = TRUE))
```

Use the mean as a threshold:  
```{r}
sabi |> 
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE)) |> 
  ggplot(aes(E, N)) +
  geom_path() +
  geom_point(aes(color = static)) + # color of points show if the points are statics or not
  coord_fixed()
```

```{r}
# filter sabi data
sabi <- sabi |> 
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE)) 

sabi_filter <- sabi |>
    filter(!static)

#plot
sabi_filter |>
    ggplot(aes(E, N)) +
    geom_path() +
    geom_point() +
    coord_fixed() +
    theme(legend.position = "bottom")
```

# Posmo data

## Preparation
```{r}
library("readr")
library("sf")

posmo <- read_delim("data/posmo_2023-01-01T00_00_00+01_00-2023-05-11T23_59_59+02_00.csv")

# Keep only the necessary columns
posmo <- select(posmo, datetime, lon_x, lat_y)
```

Transform to MN95:  
```{r}
posmo <- st_as_sf(posmo, coords = c("lon_x","lat_y"), crs = 4326) |>
  st_transform(2056)

head(posmo)
```

```{r}
posmo_coordinates <- st_coordinates(posmo)

posmo <- cbind(posmo, posmo_coordinates)
posmo
```

Select one single date:  
```{r}
posmo_filter <- posmo |>
    filter(as.Date(datetime) == "2023-05-4")
posmo_filter
```

```{r}
#plot trajectories
ggplot(posmo_filter, aes(X,Y, color = datetime)) +
  geom_point() +
  geom_path() +
  coord_equal() #so equally space North and East
```


Add useful attributes:  
```{r}
#timelag
posmo_filter <- mutate(posmo_filter, timelag = as.numeric(difftime(lead(datetime), datetime)))

#steplength
posmo_filter <- posmo_filter %>% mutate(steplength_m = sqrt((X-lead(X))^2 + (Y-lead(Y))^2))

#speed
posmo_filter <- posmo_filter %>% mutate(speed_ms = steplength_m/timelag)
```


## Task 1: Segmentation

### Step a) specify a temporal window
The timelag is of 9 seconds excepted when error of recording.  
So a temporal window of 60 minutes will give us ~6 fixes.
```{r}
#visualise:
posmo_filter %>% head(50) %>% 
  ggplot(aes(datetime, 1)) +
  geom_point()
```

Temporal window of 60 minutes: 6 fixes
pos[n-3] to pos[n]  # distance to pos -27 seconds
pos[n-2] to pos[n]  # distance to pos -18 seconds
pos[n-1] to pos[n]  # distance to pos -9 seconds
pos[n] to pos[n+1]  # distance to pos +9 seconds
pos[n] to pos[n+2]  # distance to pos +18 seconds
pos[n] to pos[n+3]  # distance to pos +27 seconds

```{r}
posmo_filter <- posmo_filter %>%
  mutate(
    n_plus1 = sqrt((lead(X)-X)^2 + (lead(Y)-Y)^2), #from n to n+1
    n_plus2 = sqrt((lead(X, 2)-X)^2 + (lead(Y, 2)-Y)^2), #from n to n+2
    n_plus3 = sqrt((lead(X, 3)-X)^2 + (lead(Y, 3)-Y)^2), #from n to n+3
    n_minus1 = sqrt((lag(X, 1)-X)^2 + (lag(Y, 1)-Y)^2), #from n to n-1
    n_minus2 = sqrt((lag(X, 2)-X)^2 + (lag(Y, 2)-Y)^2), #from n to n-2
    n_minus3 = sqrt((lag(X, 3)-X)^2 + (lag(Y, 3)-Y)^2) #from n to n-3
  )

posmo_filter
```

### Step b) Measure the distance from every point to every other point within this temporal window
Important to specify na.rm in the mean function,  otherwise na values interfer.  

```{r}
posmo_filter <- posmo_filter |> 
  rowwise() |> #group by row with rowwise(), otherwise the mutates give one result for everything
  mutate(
    stepMean = mean(c(n_minus1, n_minus2, n_minus3, n_plus1, n_plus2, n_plus3), na.rm = TRUE) 
  ) |> 
  ungroup() #to ungroup rowwise()
```

## Task 2: Specify and apply threshold d
Explore stepMean using summary statistics:  
```{r}
summary(posmo_filter$stepMean)
```

### Step c) Remove “static points"
Determines a threshold where we consider we are not moving.  

Check if we could use the mean as the threshold:  
```{r}
ggplot(posmo_filter, aes(stepMean)) +
  geom_histogram(binwidth = 10) +
  geom_vline(xintercept = mean(posmo_filter$stepMean, na.rm = TRUE)) +
  geom_vline(xintercept = median(posmo_filter$stepMean, na.rm = TRUE), color='blue')
  
```

It seems the mean `r mean(posmo_filter$stepMean, na.rm = TRUE)` neither the median  `r median(posmo_filter$stepMean, na.rm = TRUE)` are really adequate.


## Task 3: Visualize segmented trajectories

```{r}
posmo_filter <- posmo_filter |> 
  mutate(static = stepMean < mean(posmo_filter$stepMean, na.rm = TRUE))

posmo_filter %>% 
  ggplot(aes(X, Y)) +
  geom_path() +
  geom_point(aes(color = static)) + # color of points show if the points are statics or not
  coord_fixed()
```

I am not satisfied by this result as a lot of trajectories have also been removed, but after a few trials, it seems lower threshold do not permit to remove entirely the static data.  

## Task 4: Segment-based analysis
Assign unic ids:  
```{r}
rle_id <- function(vec) {
    x <- rle(vec)$lengths
    as.factor(rep(seq_along(x), times = x))
}
```

```{r}
posmo_filter <- posmo_filter |>
    mutate(segment_id = rle_id(static))

head(posmo_filter)
```

Plot:  
```{r}
posmo_filter %>% 
  ggplot(aes(X, Y)) +
  geom_path() +
  geom_point(aes(color = segment_id)) +
  coord_fixed()
```

## Task 5: Similarity measures
Import csv:  
```{r}
pedestrian <- read_csv("data/pedestrian.csv")
pedestrian
```

```{r}

pedestrian %>% 
  ggplot(aes(E, N)) +
  geom_path() +
  geom_point(aes(color = TrajID)) +
  coord_fixed() +
  facet_wrap("TrajID")
```

The most similar trajectories seem to be:  1 and 6.
And the most dissimilar seem to be 4 and 5.

## Task 6: Calculate similarity
```{r}
library("SimilarityMeasures")
traj1 <- pedestrian |> filter(TrajID == 1)

#matrix traj1:
m <- nrow(traj1)
n <- 2 #2D
traj1_matrix <- matrix(NA, nrow = m, ncol = n)

traj1_matrix[, 1] <- traj1$E
traj1_matrix[, 2] <- traj1$N

#matrix traj6:
traj6_matrix <- matrix(NA, nrow = m, ncol = n)

traj6_matrix[, 1] <- traj1$E
traj6_matrix[, 2] <- traj1$N


```

```{r}
#DTW
DTW(traj1_matrix, traj6_matrix)
```

